### 2014-10-26 report/cover.tex
report template

\newcommand{\logo}[0]{\vspace*{1.5cm}\begin{center}\includegraphics[scale=0.4,keepaspectratio=true]{images/logo/EPFL.pdf} \end{center}\hrule\vspace{1cm}}
\linespread{1.125}\selectfont
\newcommand{\project}[1]{\begin{center}\large{#1}\end{center}}
\renewcommand{\title}[1]{\vspace{0.01cm}\begin{center}\LARGE{#1}\end{center}\vspace{0.2cm}}
\renewcommand{\author}[1]{\begin{center}\Large{#1}\end{center}}
\newcommand{\department}[1]{\begin{center}\large{#1}\end{center}}
\renewcommand{\date}[2]{\begin{center}\normalsize{#1 #2}\end{center}}

\newcommand{\supervisor}[4]{\begin{center}\begin{normalsize}{\bf #1}\\#2\\#3\\#4\end{normalsize}\end{center}}

\thispagestyle{empty}

\logo

\project{Pattern Classification and Machine Learning}
\title{Project 1}
\author{Kevin Serrano, SC-MA3\\Youssef El Baba, SC-MA3\\ Alexandre Helfre, SC-MA3}
\department{École Polytechnique Fédérale de Lausanne}
\date{November 3, 2014}

\begin{center}
\begin{tabular}{cc}
\begin{tabular}{p{6cm}}
\supervisor{Professor}{Mohammad Emtiyaz Khan}{LCA4}{EPFL}

\end{tabular}
\end{tabular}
\end{center}

\clearpage \thispagestyle{empty} \cleardoublepage