### 2014-11-27 Project 2/hw5/hw5.tex
hw5 extracted

\documentclass[11pt,a4paper,notitlepage]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{enumerate}
\usepackage[left=2.00cm, right=2.00cm, top=2.00cm, bottom=2.00cm]{geometry}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{arydshln}
\lhead{Big Data}

\chead{Homework 5}
\rhead{Kevin Serrano, 204141}
\pagestyle{fancy}
\author{Kevin Serrano}
\title{Homework 5}

\begin{document}
\maketitle
\section*{Part 1}
If we multiply $R$ by $R^t$, we obtain a similarity matrix. Row $i$ shows the number of items that are liked by user $i$ and $j$, where the user $j$ is the $j$-th column.\\
\underline{Example:}\\
Let $m = 3$ and $n = 5$. Define $$R = \left[\begin{matrix}
0 & 1 &    0  &   1 &    1 \\
     1  &   0   &  0 &     1  &   0\\
     0   &  1   &  1  &   1   &  0
\end{matrix}\right]$$ Then $$R \cdot R^t = \left[ \begin{matrix}
     3  &   1   &  2\\
     1  &   2  &   1\\
     2  &   1 &    3
\end{matrix}\right]$$
That means user $1$ likes $3$ items in common with himself (obviously), 1 with user 2 and 2 with user 3. \\

By the definition of the cosine similarity, we need to divide each $(i,j)$ by the norm of the number of item liked by the user $i$ and $j$ (the norm is the square root here). To applied this to the all matrix, we just need to divide $R\cdot R^t$ by the sqrt of $P$ : $\sqrt{P}$.\\
Thus $$S_U = \sqrt{P}^{-1} \cdot R \cdot R^t \cdot \sqrt{P}^{-1}$$
For the example we obtain $$\frac{R \cdot R^t}{\sqrt{P} \cdot \sqrt{P}} =\sqrt{P}^{-1} \cdot R\cdot R^t\cdot  \sqrt{P}^{-1}= \left[\begin{matrix}
    1.0000  &  0.4082 &   0.6667\\
    0.4082  &  1.0000 &   0.4082\\
    0.6667  &  0.4082 &   1.0000
\end{matrix}\right]$$
where for example $0.4082 = \frac{1}{\sqrt{3} \cdot \sqrt{2}}$\\

By the same way, we can compute the cosine similarity for items as follow :\\
$R^t\cdot R$ gives us the number of user who like item $i$ and $j$. With the example it is $$R^t\cdot R =\left[ \begin{matrix}
     1  &   0   &  0&     1    & 0\\
     0  &  2    & 1  &   2    & 1\\
     0  &   1   &  1  &   1  &   0\\
     1  &  2   &  1    & 3  &   1\\
     0  &   1 &    0    & 1&     1
\end{matrix}\right]$$
That means for example that 2 users like item 2 and 4.
Now we can just divide by the sqrt of $Q$ to get the cosine similarity between items $$S_I = \sqrt{Q}^{-1} \cdot R^t \cdot R \cdot \sqrt{Q}^{-1}$$
For the example we get $$\sqrt{Q}^{-1} \cdot R^t\cdot R \cdot \sqrt{Q}^{-1} = \left[\begin{matrix}

    1.0000  &       0    &     0&    0.5774    &     0\\
         0   & 1.0000   & 0.7071 &   0.8165   & 0.7071\\
         0    &0.7071  &  1.0000  &  0.5774  &       0\\
    0.5774    &0.8165 &   0.5774   & 1.0000 &   0.5774\\
         0    &0.7071&         0    &0.5774&    1.0000
\end{matrix}\right]$$
\section*{Part 2}
\begin{enumerate}[(a)]
\item  \textbf{User-user collaborative filtering}\\

We have $$X = \left[\begin{matrix}
x_{11} & x_{12} & \cdots & x_{1n}\\
x_{21} & x_{22} & \cdots & x_{2n}\\
\vdots & \vdots & \ddots & \vdots\\
x_{m1} & x_{m2} & \cdots & x_{mn}
\end{matrix}\right]$$ where $$x_{ij} = \sum_{i=1}^{m} r_{ij} \cdot su_{ij}$$ and $r_{ij} \in R, su_{ij} \in S_U$. Need to be carefull for matrix dimension, thus we get $$X = S_U \cdot R = (\sqrt{P}^{-1} \cdot R \cdot R^t \cdot \sqrt{P}^{-1}) \cdot R$$

\item \textbf{Item-item collaborative filtering}\\
For the same reason, we have $$x_{ij} = \sum_{j=1}^{n} r_{ij} \cdot si_{ij}$$ with $r_{ij} \in R, si_{ij} \in S_I$. Thus we get $$X = R \cdot S_I = R \cdot (\sqrt{Q}^{-1} \cdot R^t \cdot R \cdot \sqrt{Q}^{-1})$$
\end{enumerate}

\section*{Part 3}
Let's apply the algorithms for user Bob ($200$-th row). I used Matlab for the implementation. See the code for more details.
\begin{enumerate}[(a)]
\item With the user-user collaborative filtering, I obtain these shows with these scores\\

\begin{tabular}{c c}
TV-show & score\\
\hline
    'FOX 28 News at 10pm' & 882.8699\\ 
    'Family Guy' & 828.5152\\
    'NBC 4 at Eleven' & 780.2172\\
    '2009 NCAA Basketball Tournament'& 765.7426\\
    'Access Hollywood'& 750.3544\\
\end{tabular}

\item With the item-item collaborative filtering, I obtain these shows with these scores\\

\begin{tabular}{c c}
TV-show & score\\
\hline
    'FOX 28 News at 10pm' & 23.8006\\    
    'NBC 4 at Eleven' & 22.6289\\
    'Access Hollywood'& 22.6210 \\
    'Family Guy'& 22.6071    \\
    'Two and a Half Men'& 22.0486
\end{tabular}

\item It is easy to check the precision. For both filtering, I get a precision of 1 for $k=5$. That means Bob have watched all movies that are proposed by the user-user or item-item filtering.
\end{enumerate}
\end{document}